---
title: "exploration_datasets"
output: html_document
---

Import required library
```{r}
library(dplyr)
library(gtsummary)
library(kableExtra)
```

```{r}
dataset_path <- "/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/raw/cll_broad_2022"
```

Read in data files
```{r}
data_clinical_patient <- read.table(file.path(dataset_path,"data_clinical_patient.txt"),header=TRUE, sep="\t")
data_clinical_sample <- read.table(file.path(dataset_path, "data_clinical_sample.txt"), header=TRUE, sep="\t")
data_mutations <- read.table(file.path(dataset_path,"data_mutations.txt"),header=TRUE, sep="\t")
```

```{r}

data_sv <- read.delim(file.path(dataset_path, "data_sv.txt"))
colnames(data_sv)
```


# data_armlevel_cna

cancer_study_identifier: cll_broad_2022
genetic_alteration_type: GENERIC_ASSAY
generic_assay_type: ARMLEVEL_CNA
datatype: CATEGORICAL
stable_id: armlevel_cna
profile_name: Putative arm-level copy-number from GISTIC
profile_description: Putative arm-level copy-number from GISTIC.
data_filename: data_armlevel_cna.txt
show_profile_in_analysis_tab: true
generic_entity_meta_properties: NAME,DESCRIPTION

```{r}
data_armlevel_cna   <- read.table(file.path(dataset_path,"data_armlevel_cna.txt"),header=TRUE, sep="\t") 
head(colnames(data_armlevel_cna))
unique(data_armlevel_cna$ENTITY_STABLE_ID)
```

# Data_timeline_diagnosis
```{r}
data_timeline_diagnosis  <- read.table(file.path(dataset_path,"data_timeline_diagnosis.txt"),header=TRUE, sep="\t")                 
colnames(data_timeline_diagnosis)
```

#data_timeline_specimen  
```{r}
data_timeline_treatment  <- read.table(file.path(dataset_path,"data_timeline_treatment.txt"),header=TRUE, sep="\t") 
colnames(data_timeline_treatment)
```


```{r}
data_timeline_prior_treatment  <- read.table(file.path(dataset_path,"data_timeline_prior_treatment.txt"),header=TRUE, sep="\t")           

```
```{r}

data_timeline_death  <- read.table(file.path(dataset_path,"data_timeline_death.txt"),header=TRUE, sep="\t")       


data_clinical_patient   <- read.table(file.path(dataset_path,"data_clinical_patient.txt"),header=TRUE, sep="\t") 


data_clinical_sample          <- read.table(file.path(dataset_path,"data_clinical_sample.txt"),header=TRUE, sep="\t")       


data_focal_cna         <- read.table(file.path(dataset_path,"data_focal_cna.txt"),header=TRUE, sep="\t")               

data_timeline_specimen  <- read.table(file.path(dataset_path,"data_timeline_specimen.txt"),header=TRUE, sep="\t")                 


data_gene_panel_matrix  <- read.table(file.path(dataset_path,"data_gene_panel_matrix.txt"),header=TRUE, sep="\t")                  



data_mrna_seq_tpm_zscores_ref_all_samples<- read.table(file.path(dataset_path,"data_mrna_seq_tpm_zscores_ref_all_samples.txt"),header=TRUE, sep="\t")
head(data_mrna_seq_tpm_zscores_ref_all_samples)

meta_mutations<- read.table(file.path(dataset_path,"meta_mutations.txt"),header=TRUE, sep="\t")

data_mrna_seq_tpm<- read.table(file.path(dataset_path,"data_mrna_seq_tpm.txt"),header=TRUE, sep="\t")
head(data_mrna_seq_tpm)

meta_study<- read.table(file.path(dataset_path,"meta_study.txt"),header=TRUE, sep="\t")
data_mutations<- read.table(file.path(dataset_path,"data_mutations.txt"),header=TRUE, sep="\t")
meta_armlevel_cna  <- read.table(file.path(dataset_path,"meta_armlevel_cna.txt"),header=TRUE, sep="\t")                       
meta_sv<- read.table(file.path(dataset_path,"meta_sv.txt"),header=TRUE, sep="\t")
```
