---
title: "exploration_datasets"
output:
  pdf_document: default
  html_document: default
---

# Exploratory Data Analysis of Datasets used

Import required library
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(gtsummary)
library(kableExtra)
library(ggplot2)
```

Global path variable to the data set 
```{r}
dataset_path <- "/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/raw/cll_broad_2022"
```

Read in data files
```{r}
data_clinical_patient <- read.table(file.path(dataset_path,"data_clinical_patient.txt"),header=TRUE, sep="\t")
data_clinical_sample <- read.table(file.path(dataset_path, "data_clinical_sample.txt"), header=TRUE, sep="\t")
data_mutations <- read.table(file.path(dataset_path,"data_mutations.txt"),header=TRUE, sep="\t")
```

Number of patients with MBL and CLL respectively - data_clinical_sample
```{r}
ggplot(data_clinical_sample, aes(x=DISEASE_TYPE)) +
         geom_bar() + 
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, size=3.5) + 
  # using a color palette 
  scale_fill_viridis_d(option="mako", begin =0.2, end=0.8) + 
  theme_minimal(base_size = 14) + 
  labs(
    title = "Distribution of Disease Types",
    subtitle = "Sorted by prevalence in clinical samples",
    x = NULL, # Removes the redundant 'DISEASE_TYPE' label
    y = "Number of Cases"
  ) +
  # Clean up the grid and rotate labels if they are long
  theme(
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


library(ggplot2)
library(forcats) # This is key for the sorting "magic"

ggplot(data_clinical_sample, aes(y = fct_rev(fct_infreq(DISEASE_TYPE)), fill = DISEASE_TYPE)) +
  geom_bar(show.legend = FALSE) +
  # Adds the actual numbers inside/at the end of the bars
  geom_text(stat = 'count', aes(label = after_stat(count)), hjust = -0.2, size = 4) +
  # A sleek, professional color palette
  scale_fill_
  scale_fill_viridis_d(option = "mako", begin = 0.3, end = 0.7) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Disease Type Distribution",
    subtitle = "Sorted by frequency",
    x = "Number of Cases",
    y = NULL
  ) +
  # Extra clean-up
  theme(panel.grid.minor = element_blank())
  
  
  
library(ggplot2)

ggplot(data_clinical_sample, aes(x = DISEASE_TYPE)) +
  geom_bar(fill = "#4C78A8", width = 0.65) +
  labs(
    title = "Disease Distribution",
    x = "Disease Type",
    y = "Count"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )


ggplot(df, aes(x = reorder(disease, -..count..))) +
  geom_bar(fill = "#2A9D8F", width = 0.65) +
  labs(
    title = "Disease Cohort Composition",
    x = NULL,
    y = "Number of Patients"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.x = element_blank()
  )

```



May not use this for this stage of the project. 
Important file, should come back to it later. 
```{r}
data_sv <- read.delim(file.path(dataset_path, "data_sv.txt"))
colnames(data_sv)
```


# data_armlevel_cna - will be using this one

cancer_study_identifier: cll_broad_2022
genetic_alteration_type: GENERIC_ASSAY
generic_assay_type: ARMLEVEL_CNA
datatype: CATEGORICAL
stable_id: armlevel_cna
profile_name: Putative arm-level copy-number from GISTIC
profile_description: Putative arm-level copy-number from GISTIC.
data_filename: data_armlevel_cna.txt
show_profile_in_analysis_tab: true
generic_entity_meta_properties: NAME,DESCRIPTION

```{r}
data_armlevel_cna   <- read.table(file.path(dataset_path,"data_armlevel_cna.txt"),header=TRUE, sep="\t") 
head(colnames(data_armlevel_cna))
unique(data_armlevel_cna$ENTITY_STABLE_ID)
head(data_armlevel_cna)

```
I will be using arm level cna for now. 
```{r}
data_focal_cna <- read.table(file.path(dataset_path,"data_focal_cna.txt"),header=TRUE, sep="\t")     
colnames(data_focal_cna)
head(data_focal_cna)

```

Conclusion : I will be using arm level cna 
link : https://chatgpt.com/share/69bec9a3-62ec-8011-9869-7d6b2c0aa5f0

# Data_timeline_diagnosis
```{r}
data_timeline_diagnosis  <- read.table(file.path(dataset_path,"data_timeline_diagnosis.txt"),header=TRUE, sep="\t")                 
colnames(data_timeline_diagnosis)
```

#data_timeline_specimen  
```{r}
data_timeline_treatment  <- read.table(file.path(dataset_path,"data_timeline_treatment.txt"),header=TRUE, sep="\t") 
colnames(data_timeline_treatment)
```


```{r}
data_timeline_prior_treatment  <- read.table(file.path(dataset_path,"data_timeline_prior_treatment.txt"),header=TRUE, sep="\t")           
colnames(data_timeline_prior_treatment)
```

Use mrna seq tpm as well. it shows the level at which the genes are expressed 
```{r}
data_mrna_seq_tpm<- read.table(file.path(dataset_path,"data_mrna_seq_tpm.txt"),header=TRUE, sep="\t")
head(data_mrna_seq_tpm)
colnames(data_mrna_seq_tpm)
```


```{r}
data_timeline_death  <- read.table(file.path(dataset_path,"data_timeline_death.txt"),header=TRUE, sep="\t")       
colnames(data_timeline_death)

data_timeline_specimen  <- read.table(file.path(dataset_path,"data_timeline_specimen.txt"),header=TRUE, sep="\t")      
```


