---
title: "visualize_missing_data"
output:
  pdf_document: default
  html_document: default
---


```{r}
data <- read.table("/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/importance_score/pat_cna/patient_file_cna_rai.txt", header=TRUE, sep=" ")
```


```{r}
library(VIM)
par(cex=0.7)
aggr(data)
```



```{r}
#par(cex = 0.8)  # slightly bigger text
#par(mar = c(10, 4, 4, 2) + 0.1, cex = 0.8)

# 2. Set margins: c(bottom, left, top, right)
par(mar = c(12, 5, 2, 2), cex = 0.8)
#aggr(data,
#     col = c("skyblue", "red"),     # blue = observed, red = missing
#     numbers = TRUE,                # show counts
#     sortVars = TRUE,               # sort by missingness
#     labels = names(data),          # variable names
#     cex.axis = 0.7,                # axis text size
#     gap = 3,                       # spacing between bars
#     ylab = c("Missing data", "Pattern"))


aggr(data,
     col = c("skyblue", "red"),
     numbers = TRUE,
     sortVars = TRUE,
     labels = names(data),
     cex.axis = 0.6,   # Smaller text helps labels fit
     gap = 2,          # Reduced gap can help squeeze more in
     ylab = c("Missing Data", "Pattern"))
```
Various methods we explored to show the all column names in the x-axis. 



```{r}
library(naniar)

gg_miss_var(data)   

```


  We can see that status of treatment after sampling has the most number of missing values followed by variable having to do with treatment status and rai stage sampling which could've been done during sampling. 
  
  Imputation of such variables is not recommended because some patients might've opted for new medical advice or might've decided to not to proceed with recommended procedures. 









