---
title: "HW2 Report"
author: "Soren Basnet"
date: "`r Sys.Date()`"
output: html_document
---
# Soren Basnet
# sbasnet3@wisc.edu

```{r}
rm(list=ls())
```

Required packages/libraries
```{r}
require("FITSio")
```

Read cB58 spectrum file
```{r}
cB58 = readFrameFromFITS("cB58_Lyman_break.fit")
cB58_standardized <- (cB58$FLUX - mean(cB58$FLUX)) / sd(cB58$FLUX)
n = length(cB58_standardized)
```

Read the files from the data folder
```{r}
files = list.files("data")
# Initialize results which will later turn into hw2.csv
results = data.frame(distance=numeric(),spectrumID=character(),shift=integer(),stringsAsFactors=FALSE)
```

# Method used :
I tried distance formula, l2 minimization and landed with extraction of residuals using lm() function as the best method. 

This method made the most sense to me as I saw the ease of letting lm() do the heavy lifting while I extract the residuals. This comes at heavy computational expense as the loop takes a while to finish. 

Difficulties : Apart from the methodology, R was still difficult to use from the terminal as I have always used RStudio. 

```{r}
for (file in files) {

    noisy <- readFrameFromFITS(paste0("data/", file))
    stn_noisy <- (noisy$flux - mean(noisy$flux)) / sd(noisy$flux)
    l <- length(stn_noisy)

    possible_shifts <- 0:(l - n)

    all_distance <- numeric(length(possible_shifts))

     for (i in 1:length(possible_shifts)) {

        shift <- possible_shifts[i]
        segment <- stn_noisy[(shift + 1):(shift + n)]

        m1 <- lm(cB58_standardized ~ segment)

        all_distance[i] <- sum((m1$residuals)^2)
   }


    best_index <- which.min(all_distance)


    result = data.frame( distance = all_distance[best_index],spectrumID = file, shift = possible_shifts[best_index],
                         stringsAsFactors = FALSE )
    results <- rbind(results, result)
}

# Increasing order 
results = results[order(results$distance),]
```

Write the output to hw2.csv
```{r}
write.table(results,file="hw2.csv",sep=",",row.names=FALSE,col.names=FALSE,quote=FALSE)
```

Visualization of top 3 hits from results.

```{r}
top3 <- results[1:3, ]

for (i in 1:3) {

 file <- top3$spectrumID[i]
  shift_val <- top3$shift[i]
  
  noisy <- readFrameFromFITS(paste0("data/", file))
  stn_noisy <- (noisy$flux - mean(noisy$flux)) / sd(noisy$flux)
  
  plot(stn_noisy, type="l", col="blue", main=paste("cB58 vs", file),
       xlab="Index",
       ylab="Standardized Flux")
  
  lines(x = shift_val + (1:n),
        y = cB58_standardized,
        col = "red")
  
  legend("topright", legend = c(file, "cB58"), col = c("blue", "red"), lty = 1)
}

```

