---
title: "plots"
output: html_document
---

```{r}
z <- read.table("/Users/sorenbasnet/Desktop/STAT_443_SP_2026/guide_dataset/datafiles/fat.dat",header=FALSE)
names(z) <- c("Case","BodyfatB","BodyfatS","Density","Age","Weight","Height",
"Adiposity","FatFreewt","Neck","Chest","Abdomen","Hip","Thigh",
"Knee","Ankle","Biceps","Forearm","Wrist")
excl <- c("Case","BodyfatS","Density","Adiposity","FatFreewt")
k <- which(names(z) %in% excl)
z <- z[,-k]
```


```{r}
y <- z$BodyfatB
zc <- read.table("/Users/sorenbasnet/Desktop/STAT_443_SP_2026/guide_dataset/datafiles/cons.fit",header=TRUE) 
zc.pred <- zc$predicted
zc.res <- y-zc.pred
```


```{r}
zs <- read.table("/Users/sorenbasnet/Desktop/STAT_443_SP_2026/guide_dataset/datafiles/step.fit",header=TRUE) 
zs.pred <- zs$predicted
zs.res <- y-zs.pred
```


```{r}
zf <- read.table("/Users/sorenbasnet/Desktop/STAT_443_SP_2026/guide_dataset/datafiles/gf.fit",header=TRUE) 
zf.pred <- zf$predicted
zf.res <- y-zf.pred
```

```{r}
library(rpart)
rp <- rpart(BodyfatB ~ .,data=z, method="anova") ## rpart
rp.pred <- predict(rp,data=z)
rp.res <- y - rp.pred
```


```{r}
library(randomForest)
rf <- randomForest(BodyfatB ~ .,data=z) ## randomForest
rf.pred <- predict(rf,data=z)
rf.res <- y - rf.pred
```


```{r}
library(ranger)
ran <- ranger(BodyfatB ~ .,data=z) ## ranger
ran.pred <- predict(ran,data=z)$predictions
ran.res <- y - ran.pred
```



```{r}
### predicted vs observed
all <- c(y,zc.pred,zs.pred,zf.pred,rp.pred,rf.pred,ran.pred)
par(mfrow=c(2,3))
plot(y,zc.pred,xlab="Observed",ylab="Predicted",main="Guide constant",ylim=range(all))
abline(c(0,1),col="red")
plot(y,zs.pred,xlab="Observed",ylab="Predicted",main="Guide stepwise",ylim=range(all))
abline(c(0,1),col="red")
plot(y,zf.pred,xlab="Observed",ylab="Predicted",main="Guide forest",ylim=range(all))
abline(c(0,1),col="red")
plot(y,rp.pred,xlab="Observed",ylab="Predicted",main="rpart",ylim=range(all))
abline(c(0,1),col="red")
plot(y,rf.pred,xlab="Observed",ylab="Predicted",main="randomForest",ylim=range(all))
abline(c(0,1),col="red")
plot(y,ran.pred,xlab="Observed",ylab="Predicted",main="ranger",ylim=range(all))
abline(c(0,1),col="red")
```


```{r}
### residuals vs predicted
all.res <- c(zc.res,zs.res,zf.res,rp.res,rf.res,ran.res)
par(mfrow=c(2,3))
plot(zc.res ~ zc.pred,xlab="Predicted",ylab="Residuals",main="Guide constant",
ylim=range(all.res))
abline(h=0,col="red")
plot(zs.res ~ zs.pred,xlab="Predicted",ylab="Residuals",main="Guide stepwise",
ylim=range(all.res))
abline(h=0,col="red")
plot(zf.res ~ zf.pred,xlab="Predicted",ylab="Residuals",main="Guide forest",
ylim=range(all.res))
abline(h=0,col="red")
plot(rp.res ~ rp.pred,xlab="Predicted",ylab="Residuals",main="rpart",
ylim=range(all.res))
abline(h=0,col="red")
plot(rf.res ~ rf.pred,xlab="Predicted",ylab="Residuals",main="randomForest",
ylim=range(all.res))
abline(h=0,col="red")
plot(ran.res ~ ran.pred,xlab="Predicted",ylab="Residuals",main="ranger",
ylim=range(all.res))
abline(h=0,col="red")
```


## R code for simulation
```{r}
library(rpart)
library(ranger)
library(randomForest)
library(party)
```

```{r}


z <- read.table("/Users/sorenbasnet/Desktop/STAT_443_SP_2026/plots/realdata.txt",header=TRUE)
colnames(z)
y <- z$y 
n <- nrow(z)
nsim <- 100
ct.err <- 0 # going to get me sum of xi
ct.err2 <- 0 # going to get is sum of x bar 
rp.err <- 0
rp.err2 <- 0
rf.err <- 0
rf.err2 <- 0
ran.err <- 0
ran.err2 <- 0
gcon.err <- 0
gcon.err2 <- 0
gstep.err <- 0
gstep.err2 <- 0

```

# Compare r part, ranger, random forest and then guide 
```{r}
for(i in 1:nsim){
ran <- rbinom(n,1,0.5)
train <- ran == 1
test <- ran == 0

### ctree
ct <- ctree(y ~ .,data=z[train,]) # y here is the z$y, only give the training data to c tree 
ct.pred <- predict(ct,newdata=z[test,]) # when you predict here, you do it on the test data 
error <- mean((y[test]-ct.pred)^2) # mse 
ct.err <- ct.err+error # add it over and over again over many times 
ct.err2 <- ct.err2+error*error # accumulating squares 

### rpart
rp <- rpart(y ~ .,data=z[train,], method="anova")
rp.pred <- predict(rp,newdata=z[test,])
error <- mean((y[test]-rp.pred)^2)
rp.err <- rp.err+error
rp.err2 <- rp.err2+error*error

### randomforest
rf <- randomForest(y ~ .,data=z[train,], method="anova")
rf.pred <- predict(rf,newdata=z[test,])
error <- mean((y[test]-rf.pred)^2)
rf.err <- rf.err+error
rf.err2 <- rf.err2+error*error

### ranger
ran <- ranger(y ~ .,data=z[train,])
ran.pred <- predict(ran,data=z[test,])$predictions
error <- mean((y[test]-ran.pred)^2)
ran.err <- ran.err+error
ran.err2 <- ran.err2+error*error

### guide constant tree
z0 <- z
z0$y[test] <- NA # change the train to test - i changed it, original had train in it 
write.table(z0,"simdata.txt",row.names=FALSE,col.names=TRUE)
## IMPORTANT: choose mac or windows
system("../guide < simcons.in > simcons.log") ## mac osx
# can we run python to inside system command. if so then we can compare against python as well. 
z1 <- read.table("simcons.fit",header=TRUE) # simcons.fit is a file inside simstep.in
gcon.pred <- z1$predicted[test]
error <- mean((y[test]-gcon.pred)^2)
gcon.err <- gcon.err+error
gcon.err2 <- gcon.err2+error*error

### guide stepwise linear tree
## IMPORTANT: choose mac or windows
system("../guide < simstep.in > simstep.log") ## mac osx
z1 <- read.table("simstep.fit",header=TRUE)
gstep.pred <- z1$predicted[test]
error <- mean((y[test]-gstep.pred)^2)
gstep.err <- gstep.err+error
gstep.err2 <- gstep.err2+error*error

if(i == nsim | i==50*floor(i/50)){ # it is going to print every 5 iterations. in order to make into 2 iterations, change both iterations to 2
print(paste("Number of iterations completed =",i),quote=FALSE)
ct.mean <- ct.err/i # this is x_bar
ct.se <- sqrt((ct.err2-i*ct.mean^2)/(i*(i-1))) # this is the equation * from notes - 24 march, tuesday notes
print(paste("ctree MSE and SE =",ct.mean,ct.se),quote=FALSE) 
rp.mean <- rp.err/i
rp.se <- sqrt((rp.err2-i*rp.mean^2)/(i*(i-1)))
print(paste("rpart MSE and SE =",rp.mean,rp.se),quote=FALSE)
rf.mean <- rf.err/i
rf.se <- sqrt((rf.err2-i*rf.mean^2)/(i*(i-1)))
print(paste("randomForest MSE and SE =",rf.mean,rf.se),quote=FALSE)
ran.mean <- ran.err/i
ran.se <- sqrt((ran.err2-i*ran.mean^2)/(i*(i-1)))
print(paste("ranger MSE and SE =",ran.mean,ran.se),quote=FALSE)
gcon.mean <- gcon.err/i
gcon.se <- sqrt((gcon.err2-i*gcon.mean^2)/(i*(i-1)))
print(paste("Guide constant MSE and SE =",gcon.mean,gcon.se),quote=FALSE)
gstep.mean <- gstep.err/i
gstep.se <- sqrt((gstep.err2-i*gstep.mean^2)/(i*(i-1)))
print(paste("Guide stepwise MSE and SE =",gstep.mean,gstep.se),quote=FALSE)
}
}
```


For each row there are two numbers 
for guide step wise
17.9458853528149 - MSE
0.460701246041341 - error rate 

so mse is 17.9458853528149 +- 0.460701246041341


Ctree MSE 
29.4128019630374 +- 1.8171578022007

ranger MSE : 
23.7546154720688 +- 1.8171578022007 ( we are getting the confidence interval)


here we see that GUIDE stepwise is statistically significantly better. 

important to not that stepwise was something he came up with recently. 

L1 is a linear combination of your weight, abdomen and wrist. 

24.37 is an average. 

terminal 3 is a linear model. 
so we can tell you why your prediction was that error. 

we can put y hat produced by ai in real data.txt and then use guide to show what the ai is making decisions on 











