---
title: "cna_km_plots"
output: html_document
---
---
title: "km_plots"
output: html_document
---

Import required library
```{r}
library(survival)
library(survminer)
```

Load datafile 
```{r}
data <- read.table("/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/data_patient_file_cna_rai/patient_file_cna_rai.txt", header=TRUE, sep=" ")

head(data)
colnames(data)
summary(data$OS_MONTHS)
summary(data$surv_5yrs)
summary(data$death_5yr)
```

Create survival object 
```{r}
fit <- survfit(Surv(surv_5yrs, death_5yr) ~ data$RAI_AT_SAMPLING, data=data)

ggsurvplot(
  fit,
  data = data,
  pval = TRUE,              # shows log-rank test
  risk.table = TRUE,        # number at risk table
  conf.int = TRUE,          # confidence intervals
  xlab = "Time (months)",
  ylab = "Overall Survival Probability",
  legend.title = "Rai Stage"
)
```

This Kaplan-Meier plot illustrates that the Rai Stage at the time of sampling has no statistically significant impact on Overall Survival Probability over a 60-month period, as evidenced by the high p-value of 0.61. Across all five stages (0–4), the survival rates remain remarkably high and stable, with all groups maintaining a survival probability above approximately 75% at the five-year mark. The extensive overlap of the confidence intervals and the staggered "step" declines across the different colored cohorts further confirm that, for this specific study population, clinical staging did not serve as a primary predictor of mortality outcomes within the observed timeframe.

```{r}
coxph(Surv(surv_5yrs, death_5yr) ~ X17p_status, data = data)
```


```{r}
coxph(Surv(surv_5yrs, death_5yr) ~ X17p_status + AGE + SEX, data = data)
```





# Making curves for coxph(Surv(surv_5yrs, death_5yr) ~ X17p_status + AGE + SEX, data=data)


```{r}
library(survminer)

# 1. To create the Forest Plot
model <- coxph(Surv(surv_5yrs, death_5yr) ~ X17p_status + AGE + SEX, data = data)
ggforest(model, data = data)

# 2. To create Adjusted Survival Curves by Rai Stage
ggadjustedcurves(model, variable = "X17p_status", data = data)

```










