---
title: "quantile_regression"
output: html_document
---

Action : 
1) To go quantile regression on the patient data to see who survives and who does [CURRENT TASK]
    - This objective might also help with the the infection rate 
2) Then if that is successful, do quantile regression on the patient data along with genetic mutation matrix [PLANNED]

Role,Variables from your list
d (Dependent),FFS_MONTHS
c (Categorical),"CLL_EPITYPE, TUMOR_MOLECULAR_SUBTYPE, IGHV_MUTATION_STATUS, U1_STATUS, IGLV3_21_R110, SEX"
n (Numerical),"AGE, TMB_NONSYNONYMOUS, TUMOR_SAMPLE_PURITY, IGHV_IDENTITY_PERCENTAGE"
x (Exclude),"PATIENT_ID, SAMPLE_ID, OS_STATUS, OS_MONTHS (to avoid leakage), CENTER"
s (Split-only),Any genetic mutation columns (0/1) from your third file.

Import required library

```{r}

library(dplyr)
library(gtsummary)
library(kableExtra)

```

Import required datasets

```{r}

patient_file <- read.table("/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/processed/patient_file.txt", header = TRUE, sep=" ")
mutation_matrix <- read.table("/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/interim/mutation_matrix_march_7.txt", header=TRUE, sep=" ")

```

```{r}

dim(patient_file)
dim(mutation_matrix)

```


```{r}

colnames(patient_file)

```

```{r}

patient_mutation <- left_join(patient_file, mutation_matrix, by = c("SAMPLE_ID" = "Tumor_Sample_Barcode"))
dim(patient_mutation)

```

Save the patient_mutation file 

```{r}

write.table(patient_mutation, "/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/processed/data_patient_mutation_march_8.txt", sep=" ", row.names = FALSE, quote = FALSE)

```




