---
title: "01_variable_setup"
output: html_document
---

Import required library
```{r}
library(dplyr)
```

Import required dataset(s)
```{r}
dataset_path <- "/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/raw/cll_broad_2022"
```

Read in data files
```{r}
data_clinical_patient <- read.table(file.path(dataset_path,"data_clinical_patient.txt"),header=TRUE, sep="\t")
data_clinical_sample <- read.table(file.path(dataset_path, "data_clinical_sample.txt"), header=TRUE, sep="\t")
```

Number of patients in the data_clinical_patient file
```{r}
length(unique(data_clinical_patient$PATIENT_ID))
nrow(data_clinical_patient)
```

Lets see what patient ID are getting repeated 
```{r}
table(data_clinical_sample$PATIENT_ID)[table(data_clinical_sample$PATIENT_ID) > 1]

dup_ids <- data_clinical_sample$PATIENT_ID[duplicated(data_clinical_sample$PATIENT_ID)]
data_dups <- data_clinical_sample[data_clinical_sample$PATIENT_ID %in% dup_ids, ]
```

We will only keep the first occurence 
```{r}
data_unique <- data_clinical_sample[!duplicated(data_clinical_sample$PATIENT_ID), ]
```


Number of patients in the data_clinical_sample file
```{r}
length(unique(data_clinical_sample$PATIENT_ID))
nrow(data_clinical_sample)
```

Column names 

```{r}
colnames(data_clinical_patient)
```

```{r}
colnames(data_unique)
```

Merge the 2 patient files
```{r}
patient_file <- left_join(data_unique, data_clinical_patient, by = "PATIENT_ID")
length(unique(patient_file$PATIENT_ID))
nrow(patient_file)
```


```{r}
data_armlevel_cna   <- read.table(file.path(dataset_path,"data_armlevel_cna.txt"),header=TRUE, sep="\t") 
head(colnames(data_armlevel_cna))
print("---------------------------------------")
unique(data_armlevel_cna$ENTITY_STABLE_ID)
print("---------------------------------------")
dim(data_armlevel_cna)
```
Change the column names to something suitable for merge 
```{r}
colnames(data_armlevel_cna) <- gsub("\\.", "-", colnames(data_armlevel_cna))
head(colnames(data_armlevel_cna))
```

Add colnames - 11q_status, 17p_status, 17q_status, etc to the patient file 
```{r}
cna_mat <- data_armlevel_cna[, -(1:3)]
cna_t <- as.data.frame(t(cna_mat))
colnames(cna_t) <- data_armlevel_cna$ENTITY_STABLE_ID
cna_t$SAMPLE_ID <- rownames(cna_t)
rownames(cna_t) <- NULL
length(unique(cna_t$SAMPLE_ID))
```


Merge the patient file with cna_t file 
```{r}
patient_file_cna <- merge(data_clinical_sample, cna_t, by = "SAMPLE_ID", all.x = TRUE)
```





