--
title: "data_clinical_patient"
output:
  pdf_document: default
  html_document: default
---

Import required library
```{r}
library(dplyr)
library(gtsummary)
library(kableExtra)
```

```{r}
dataset_path <- "/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/raw/cll_broad_2022"
```

Read in data files
```{r}
data_clinical_patient <- read.table(file.path(dataset_path,"data_clinical_patient.txt"),header=TRUE, sep="\t")
data_clinical_sample <- read.table(file.path(dataset_path, "data_clinical_sample.txt"), header=TRUE, sep="\t")
data_mutations <- read.table(file.path(dataset_path,"data_mutations.txt"),header=TRUE, sep="\t")
```

Column Names - clinical patient
```{r}
colnames(data_clinical_patient)
```

Clinical patient file clean up
```{r}
# 1. Change in data
data_clinical_patient[] <- lapply(data_clinical_patient, function(x) gsub("-", "_", x))
data_clinical_patient[] <- lapply(data_clinical_patient, function(y) gsub(" ", "", y))
data_clinical_patient[] <- lapply(data_clinical_patient, function(z) gsub(":", "_", z))

# 2. Change in column names
# names(data_clinical_patient) <- gsub(",", " ", names(data_clinical_patient))

data_clinical_patient$OS_STATUS <- as.factor(data_clinical_patient$OS_STATUS)
data_clinical_patient$COHORT <- as.factor(data_clinical_patient$COHORT)
data_clinical_patient$IGHV_MUTATION_STATUS <- as.factor(data_clinical_patient$IGHV_MUTATION_STATUS)
data_clinical_patient$TREATMENT_STATUS <- as.factor(data_clinical_patient$TREATMENT_STATUS)
data_clinical_patient$PRIOR_TREATMENT_CATEGORY <- as.factor(data_clinical_patient$PRIOR_TREATMENT_CATEGORY)
data_clinical_patient$TREATMENT_AFTER_SAMPLING <- as.factor(data_clinical_patient$TREATMENT_AFTER_SAMPLING)
data_clinical_patient$EXPRESSION_CLUSTER <- as.factor(data_clinical_patient$EXPRESSION_CLUSTER)
data_clinical_patient$FFS_STATUS <- as.factor(data_clinical_patient$FFS_STATUS)
data_clinical_patient$FFS_MONTHS <- as.numeric(data_clinical_patient$FFS_MONTHS)

data_clinical_patient$IGHV_MUTATION_STATUS[data_clinical_patient$IGHV_MUTATION_STATUS == ""] <- NA

data_clinical_patient <- data_clinical_patient %>%
  mutate(IGHV_MUTATION_STATUS = case_when(
    IGHV_MUTATION_STATUS == "0" ~ NA_character_,
    IGHV_MUTATION_STATUS == " " ~ NA_character_,
    TRUE ~ IGHV_MUTATION_STATUS
  ))

data_clinical_patient$IGHV_MUTATION_STATUS <- as.factor(data_clinical_patient$IGHV_MUTATION_STATUS)


data_clinical_patient$TREATMENT_STATUS[data_clinical_patient$TREATMENT_STATUS == ""] <- NA

data_clinical_patient <- data_clinical_patient %>%
  mutate(TREATMENT_STATUS = case_when(
    TREATMENT_STATUS == "0" ~ NA_character_,
    TREATMENT_STATUS == " " ~ NA_character_,
    TRUE ~ TREATMENT_STATUS
  ))

data_clinical_patient$TREATMENT_STATUS <- as.factor(data_clinical_patient$TREATMENT_STATUS)
data_clinical_patient$SEX <- as.factor(data_clinical_patient$SEX)



#unique(data_clinical_patient$OS_STATUS)
#unique(data_clinical_patient$COHORT)
#unique(data_clinical_patient$IGHV_MUTATION_STATUS)
#unique(data_clinical_patient$TREATMENT_STATUS) 
#unique(data_clinical_patient$PRIOR_TREATMENT_CATEGORY)
#unique(data_clinical_patient$TREATMENT_AFTER_SAMPLING) 
#unique(data_clinical_patient$EXPRESSION_CLUSTER)
#unique(data_clinical_patient$FFS_STATUS) 

data_clinical_patient$AGE <- as.numeric(data_clinical_patient$AGE)
data_clinical_patient$AGE_SAMPLING <- as.numeric(data_clinical_patient$AGE_SAMPLING)
data_clinical_patient$OS_MONTHS <- as.numeric(data_clinical_patient$OS_MONTHS)
data_clinical_patient$DEATH_DAYS <- as.numeric(data_clinical_patient$DEATH_DAYS)
data_clinical_patient$IGHV_IDENTITY_PERCENTAGE <- as.numeric(data_clinical_patient$IGHV_IDENTITY_PERCENTAGE)



summary(data_clinical_patient)
```


Write to interim data file
```{r}
write.table(data_clinical_patient, "/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/interim/data_clinical_patient_march_7.txt", sep=" ", row.names = FALSE, quote = FALSE)
```



###############################################################################################

###############################################################################################
###############################################################################################
###############################################################################################
###############################################################################################

###############################################################################################

Column names - clinical sample 
```{r}
colnames(data_clinical_sample)
```

Left join dataframe - clinical sample and clinical patient file
```{r}
patient_file <- left_join(data_clinical_sample, data_clinical_patient, by = "PATIENT_ID")

#d1 <- left_join(patient_file, data_mutations, by = c("SAMPLE_ID" = "Tumor_Sample_Barcode"))
#colnames(d1)
#length(unique(d1$PATIENT_ID))
#length(unique(d1$SAMPLE_ID))

```

Column names for patient_file 
```{r}
colnames(patient_file)
```


Save the patient file
```{r}
patient_file[] <- lapply(patient_file, function(x) {
  if (is.character(x)) gsub(",", "_", x) else x
  if (is.character(x)) gsub(" ", "_", x) else x
})

write.table(patient_file, "/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/interim/patient_file.txt", sep=" ", row.names = FALSE, quote = FALSE)
```


Merge the file with the transposed hugo data 
```{r}

mutation_matrix <- read.table("/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/interim/mutation_matrix.txt",header=TRUE, sep=" ")

mutation_summary <- mutation_matrix %>%
  group_by(Tumor_Sample_Barcode) %>%
  summarise(
    Consequence = paste(unique(Consequence), collapse = "; "),
    .groups = "drop"
  )

mutation_matrix %>%
  count(Tumor_Sample_Barcode) %>%
  filter(n > 1)


patient_mutation <- left_join(patient_file, mutation_matrix, by = c("SAMPLE_ID" = "Tumor_Sample_Barcode"))

```


Save the patient_mutation file 
```{r}
patient_mutation[] <- lapply(patient_mutation, function(x) {
  if (is.character(x)) gsub(",", "_", x) else x
  if (is.character(x)) gsub(" ", "_", x) else x
})

patient_mutation[is.na(patient_mutation)] <- NA

# Just for now please 
#patient_mutation <- patient_mutation %>% select(-CANCER_TYPE_DETAILED)

write.table(patient_mutation, "/Users/sorenbasnet/Documents/Github/STAT_443_FINAL_PROJECT/data/interim/patient_mutation.txt", sep=" ", row.names = FALSE, quote = FALSE)

patient_mutation[25]


```








